PANCHANGE_data_preprocessing

### 2020-08-04 PANCHANGE_data_preprocessing2020-07-15.Rmd
Added path for OCI-R data and variable names to the cleaning block

## COPING GLAD: GAD-7

@@

```{r PHQ baseline missingness 8 items}
#We need to identify those who have all items APART from the suicide ideation item - to still include these individuals. 
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    phq.base_cut_off = 
      if_else(
        na_per_person_phq_base == 1 & 
          is.na(phq.dead_hurting_thoughts_base_numeric), 
        1,
        0
      )
  )

table(dat.no.sex.no.gender.no.age.no.ibd$phq.base_cut_off)
```

@@

```{r PHQ retro missingness 8 items}
#We need to identify those who have all items APART from the suicide ideation item - to still include these individuals. 
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    phq.retro_cut_off = 
      if_else(
        na_per_person_phq_retro == 1 & 
          is.na(phq.dead_hurting_thoughts_retro_numeric), 
        1,
        0
      )
  )

table(dat.no.sex.no.gender.no.age.no.ibd$phq.retro_cut_off)
```

@@

```{r PHQ prepan missingness 8 items}
#We need to identify those who have all items APART from the suicide ideation item - to still include these individuals. 
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    phq.prepan_cut_off = 
      if_else(
        na_per_person_phq_prepan == 1 & 
          is.na(phq.dead_hurting_thoughts_prepan_numeric), 
        1,
        0
      )
  )

table(dat.no.sex.no.gender.no.age.no.ibd$phq.prepan_cut_off)
```

@@

#Sum scores missingness cut offs
KT: This series of chunks creates a "full" sum score for the PHQ, GAD and PCL.

PHQ
The PHQ sum score is set to NA if they have more than one missing item, and it is not the last suicide ideation item (they do not meet the cut off criteria - phq.base_cut_off)
```{r sum score cut offs PHQ base}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    phq.sum_score_base_full =
      if_else(na_per_person_phq_base > 1 & phq.base_cut_off == 0,
              NA_real_,
              phq.sum_score_base
              )
  )

summary(dat.no.sex.no.gender.no.age.no.ibd$phq.sum_score_base)
summary(dat.no.sex.no.gender.no.age.no.ibd$phq.sum_score_base_full)
```

```{r sum score cut offs PHQ prepan}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    phq.sum_score_prepan_full =
      if_else(na_per_person_phq_prepan > 1 & phq.prepan_cut_off == 0,
              NA_real_,
              phq.sum_score_prepan
              )
  )

summary(dat.no.sex.no.gender.no.age.no.ibd$phq.sum_score_prepan)
summary(dat.no.sex.no.gender.no.age.no.ibd$phq.sum_score_prepan_full)
```

```{r sum score cut offs PHQ retro}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    phq.sum_score_retro_full =
      if_else(na_per_person_phq_retro > 1 & phq.retro_cut_off == 0,
              NA_real_,
              phq.sum_score_retro
              )
  )

summary(dat.no.sex.no.gender.no.age.no.ibd$phq.sum_score_retro)
summary(dat.no.sex.no.gender.no.age.no.ibd$phq.sum_score_retro_full)
```

GAD
The GAD sum score is set to NA if they have any missing items
```{r sum score cut offs GAD base}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    gad.sum_score_base_full =
      if_else(na_per_person_gad_base > 0,
              NA_real_,
              gad.sum_score_base
              )
  )

summary(dat.no.sex.no.gender.no.age.no.ibd$gad.sum_score_base)
summary(dat.no.sex.no.gender.no.age.no.ibd$gad.sum_score_base_full)
```

```{r sum score cut offs GAD prepan}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    gad.sum_score_prepan_full =
      if_else(na_per_person_gad_prepan > 0,
              NA_real_,
              gad.sum_score_prepan
              )
  )

summary(dat.no.sex.no.gender.no.age.no.ibd$gad.sum_score_prepan)
summary(dat.no.sex.no.gender.no.age.no.ibd$gad.sum_score_prepan_full)
```

```{r sum score cut offs GAD retro}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    gad.sum_score_retro_full =
      if_else(na_per_person_gad_retro > 0,
              NA_real_,
              gad.sum_score_retro
              )
  )

summary(dat.no.sex.no.gender.no.age.no.ibd$gad.sum_score_retro)
summary(dat.no.sex.no.gender.no.age.no.ibd$gad.sum_score_retro_full)
```

PCL
The PCL sum score is set to NA if they have any missing items - only for basline and prepan.
```{r sum score cut offs PCL base}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    pcl.sum_score_base_full =
      if_else(na_per_person_pcl_base > 0,
              NA_real_,
              pcl.sum_score_base
              )
  )

summary(dat.no.sex.no.gender.no.age.no.ibd$pcl.sum_score_base)
summary(dat.no.sex.no.gender.no.age.no.ibd$pcl.sum_score_base_full)
```

```{r sum score cut offs PCL prepan}
dat.no.sex.no.gender.no.age.no.ibd <- dat.no.sex.no.gender.no.age.no.ibd %>%
  mutate(
    pcl.sum_score_prepan_full =
      if_else(na_per_person_pcl_prepan > 0,
              NA_real_,
              pcl.sum_score_prepan
              )
  )

summary(dat.no.sex.no.gender.no.age.no.ibd$pcl.sum_score_prepan)
summary(dat.no.sex.no.gender.no.age.no.ibd$pcl.sum_score_prepan_full)
```

>>>>

# glad ocir

```{r glad ocir cols}
glad.ocir.raw <- readRDS(file = paste0(data.raw_path, "/2020-08-03/glad/ocir_glad.rds"))
dim(glad.ocir.raw)
colnames(glad.ocir.raw)
```

```{r glad ocir cleaning}
glad.ocir.raw.id <- glad.ocir.raw %>%
  drop_na(externalDataReference) %>% # Drop NAs
  distinct(externalDataReference, .keep_all = TRUE) %>% # Remove duplicates based on ID
  separate(externalDataReference, into = c("Sample", "ID"), sep = 4) %>% # Split ID in Sample and Number
  mutate(ID = as.numeric(ID),
         startDate.glad = startDate) %>% # Rename startDate
  select(
    Sample, # Sample
    ID, # ID
    ocir.i_repeatedly_check_doors_windows_drawers_etc_prepan = ocir.i_repeatedly_check_doors_windows_drawers_etc.,
    ocir.arranged_things_change_upset_prepan = ocir.arranged_things_change_upset,
    ocir.repeat_feel_numbers_prepan = ocir.repeat_feel_numbers,
    ocir.clean_simply_wash_feel_prepan = ocir.clean_simply_wash_feel,
    ocir.mind_upset_unpleasant_thoughts_prepan = ocir.mind_upset_unpleasant_thoughts,
    ocir.afraid_avoid_throwing_things_prepan = ocir.afraid_avoid_throwing_things,
    ocir.light_switches_water_taps_prepan = ocir.light_switches_water_taps,
    ocir.arranged_things_prepan = ocir.arranged_things,
    ocir.good_feel_bad_numbers_prepan = ocir.good_feel_bad_numbers,
    ocir.hands_longer_wash_prepan = ocir.hands_longer_wash,
    ocir.saved_things_prepan = ocir.saved_things,
    ocir.frequently_difficulty_rid_nasty_prepan = ocir.frequently_difficulty_rid_nasty,
    ocir.i_check_things_more_often_than_necessary_prepan = ocir.i_check_things_more_often_than_necessary.,
    ocir.arranged_properly_objects_upset_prepan = ocir.arranged_properly_objects_upset,
    ocir.feel_compelled_count_things_prepan = ocir.feel_compelled_count_things,
    ocir.strangers_touch_touched_difficult_prepan = ocir.strangers_touch_touched_difficult,
    ocir.control_difficult_find_thoughts_prepan = ocir.control_difficult_find_thoughts,
    ocir.i_collect_things_i_dont_need_prepan = ocir.i_collect_things_i_dont_need
  ) %>%
  mutate_if(is.numeric, ~na_if(., -88)) %>% # Recode missing values to NAs in numeric variables
  mutate_if(is.numeric, ~na_if(., -99)) %>%
  mutate_if(is.numeric, ~na_if(., -77)) %>%
  mutate_if(is.factor, ~na_if(., "Seen but not answered")) %>% # Recode missing values to NAs in factor variables
  mutate_if(is.factor, ~na_if(., "Don't know")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to say")) %>%
  mutate_if(is.factor, ~na_if(., "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Seen but not answered")) %>% # Drop empty factor levels
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Don't know")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to answer")) %>%
  mutate_if(is.factor, ~forcats::fct_drop(., only = "Prefer not to say"))

# Inspect dimensions
dim(glad.ocir.raw.id)
# Inspect colnames
colnames(glad.ocir.raw.id)
#Differences
dim(glad.ocir.raw)[1]-dim(glad.ocir.raw.id)[1]
```

Calculate sum score for OCI-R

*** MD - need to adapt for participants with missing data (should be NA)
```{r ocir sum score prepan}
# Create vector with scoring key: If the item is reversed coded use -1. 
# There needs to be the correct number of values for each item in the questionnaire. 
ocir.items_key_prepan <- c(1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1,1)

# Create vector with items that are included in sum score
ocir.items_prepan = c(
  "ocir.all_variables_listed_here",
  ##add variables to include in sum score
  )

# Calulate sum score
ocir.scored_items_prepan <- scoreItems(keys = ocir.items_key_prepan,
                                        items = glad.ocir.raw.id[ocir.items_prepan],
                                        totals = TRUE, 
                                        missing = TRUE, 
                                        impute = 'none', 
                                        min = 0, 
                                        max = 3)
# Add in column to tibble
glad.ocir.raw.id$ocir.sum_score_prepan <- ocir.scored_items_prepan$scores

# Look at values
glad.ocir.raw.id %>%
  descr(ocir.sum_score_prepan,
        stats = "common"
        )
```

```{r glad ocir head}
head(glad.ocir.raw.id)
```

```{r glad ocir summary}
summarytools::dfSummary(
  glad.ocir.raw.id,
  graph.col = F,
  valid.col = F
  )
```

# COPING GLAD: GAD-7

@@

@@

@@

@@

#Creating "complete cases" datasets  - NOTE THIS PROBABLY WONT BE USED BUT JUST TO GIVE AN IDEA OF COMPLETE CASES NUMBERS. KT. 
Complete case data for PHQ, GAD and PCL - 3345 remaining, drop 35,036.
```{r complete cases only}
all.items <- c(phq.items_prepan,
                phq.items_base,
                phq.items_retro,
                gad.items_prepan,
                gad.items_base,
                gad.items_retro,
                pcl.items_prepan,
                pcl.items_base,
                pcl.items_retro)

dat.no.sex.no.gender.no.age.no.ibd.complete.cases <- dat.no.sex.no.gender.no.age.no.ibd[complete.cases(dat.no.sex.no.gender.no.age.no.ibd[, all.items ]),]

dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases)
dim(dat.no.sex.no.gender.no.age.no.ibd)[1]-dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases)[1]
```

PHQ
Complete case data for PHQ baseline (COPING) - 9,260 remaining
```{r complete cases phq base}
dat.no.sex.no.gender.no.age.no.ibd.complete.cases.phq.base <- dat.no.sex.no.gender.no.age.no.ibd[complete.cases(dat.no.sex.no.gender.no.age.no.ibd[, phq.items_base ]),]

dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.phq.base)
dim(dat.no.sex.no.gender.no.age.no.ibd)[1]-dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.phq.base)[1]
```

Complete case data for PHQ pre-pandemic (GLAD) - 37,308 remaining
```{r complete cases phq prepan}
dat.no.sex.no.gender.no.age.no.ibd.complete.cases.phq.prepan <- dat.no.sex.no.gender.no.age.no.ibd[complete.cases(dat.no.sex.no.gender.no.age.no.ibd[, phq.items_prepan ]),]

dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.phq.prepan)
dim(dat.no.sex.no.gender.no.age.no.ibd)[1]-dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.phq.prepan)[1]
```

Complete case data for PHQ retrospective (COPING) - 9,262 remaining
```{r complete cases phq retro}
dat.no.sex.no.gender.no.age.no.ibd.complete.cases.phq.retro <- dat.no.sex.no.gender.no.age.no.ibd[complete.cases(dat.no.sex.no.gender.no.age.no.ibd[, phq.items_retro ]),]

dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.phq.retro)
dim(dat.no.sex.no.gender.no.age.no.ibd)[1]-dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.phq.retro)[1]
```

GAD
Complete case data for GAD baseline (COPING) - 12,457 remaining
```{r complete cases gad base}
dat.no.sex.no.gender.no.age.no.ibd.complete.cases.gad.base <- dat.no.sex.no.gender.no.age.no.ibd[complete.cases(dat.no.sex.no.gender.no.age.no.ibd[, gad.items_base ]),]

dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.gad.base)
dim(dat.no.sex.no.gender.no.age.no.ibd)[1]-dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.gad.base)[1]
```

Complete case data for gad pre-pandemic (GLAD) - 36,149 remaining
```{r complete cases gad prepan}
dat.no.sex.no.gender.no.age.no.ibd.complete.cases.gad.prepan <- dat.no.sex.no.gender.no.age.no.ibd[complete.cases(dat.no.sex.no.gender.no.age.no.ibd[, gad.items_prepan ]),]

dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.gad.prepan)
dim(dat.no.sex.no.gender.no.age.no.ibd)[1]-dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.gad.prepan)[1]
```

Complete case data for gad retrospective (COPING) - 8,930 remaining
```{r complete cases gad retro}
dat.no.sex.no.gender.no.age.no.ibd.complete.cases.gad.retro <- dat.no.sex.no.gender.no.age.no.ibd[complete.cases(dat.no.sex.no.gender.no.age.no.ibd[, gad.items_retro ]),]

dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.gad.retro)
dim(dat.no.sex.no.gender.no.age.no.ibd)[1]-dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.gad.retro)[1]
```

PCL
Complete case data for PCL baseline (COPING) - 12,376 remaining
```{r complete cases pcl base}
dat.no.sex.no.gender.no.age.no.ibd.complete.cases.pcl.base <- dat.no.sex.no.gender.no.age.no.ibd[complete.cases(dat.no.sex.no.gender.no.age.no.ibd[, pcl.items_base ]),]

dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.pcl.base)
dim(dat.no.sex.no.gender.no.age.no.ibd)[1]-dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.pcl.base)[1]
```

Complete case data for PCL pre-pandemic (GLAD) - 35,176 remaining
```{r complete cases pcl prepan}
dat.no.sex.no.gender.no.age.no.ibd.complete.cases.pcl.prepan <- dat.no.sex.no.gender.no.age.no.ibd[complete.cases(dat.no.sex.no.gender.no.age.no.ibd[, pcl.items_prepan ]),]

dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.pcl.prepan)
dim(dat.no.sex.no.gender.no.age.no.ibd)[1]-dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.pcl.prepan)[1]
```

Complete case data for PCL retrospective (COPING) - 4,618 remaining
```{r complete cases pcl retro}
dat.no.sex.no.gender.no.age.no.ibd.complete.cases.pcl.retro <- dat.no.sex.no.gender.no.age.no.ibd[complete.cases(dat.no.sex.no.gender.no.age.no.ibd[, pcl.items_retro ]),]

dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.pcl.retro)
dim(dat.no.sex.no.gender.no.age.no.ibd)[1]-dim(dat.no.sex.no.gender.no.age.no.ibd.complete.cases.pcl.retro)[1]
```
